---
Author: "Dania Jafar"
title: "Homework 3"
output: github_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

```{r}
library(p8105.datasets)
data("instacart")
ins_aisles = instacart %>% 
count(aisle, name = "aisle_order_count") %>% 
arrange(desc(aisle_order_count))


library(ggplot2)
ins_aisles %>% 
filter(aisle_order_count > 10000) %>% 
ggplot(aes(x = aisle_order_count, y = aisle)) + 
  geom_point()

#create a table showing the most popular items in each of three categories
instacart2 = instacart %>% 
  dplyr::select(aisle_id, aisle, product_name) %>% 
filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
count(aisle, product_name) %>% 
group_by(aisle) %>% 
top_n(3)

```
# Description of the dataframe 
* The size of the dataset instacart is `r nrow(instacart)` rows and `r ncol(instacart)` columns.
* Some of the key variables are order_id and product_name
* The most number of orders are placed from the fresh vegetables aisle and fresh vegetables aisle. 
* The 3 most popular items ordered from these aisles are: 
-baking ingredients: Light brown sugar (499), pure baking soda (387), and cane sugar (336)
-dog food care: Snack Sticks Chicken & Rice Recipe Dog Treats (30), Organix Chicken & Brown Rice Recipe (28), Small Dog Biscuits (26)
-packaged food vegetables: Organic Baby Spinach (9784), Organic Raspberries (5546), Organic Blueberries (4966)

```{r Coffee Ice cream & Apples}
#A table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
icecream_apples = instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(order_hour_of_day, product_name) %>% 
  avg()

```


